---
title: "Simulating the GORP"
author: "N.R Okereke"
date: "2025-01-09"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


***Load Packages***
```{r}
library(AlphaSimR)
library(tidyverse)
```

```{r}
set.seed(7)
```

***Simulate male sterile lines***
```{r}
# Donor line
ms_lines <- runMacs2(nInd = 3, nChr = 10, segSites = 1000, genLen = 2, inbred = TRUE)
```

***Simulate Diverse Founder Lines***
```{r}
# Diverse founder lines (35 founders for the NAM population)
diverseFounders <- runMacs2(nInd = 35, nChr = 10, segSites = 1000, genLen = 2, inbred = TRUE)
```

***Simulate parameters for both donor and diverse lines***
```{r}
# SimParam for reference and diverse founders
SP <- SimParam$new(ms_lines)
SP$addTraitA(nQtlPerChr = 5)
SP$setSexes(sexes = "yes_rand")

SP2 <-SimParam$new(diverseFounders)
SP2$addTraitA(nQtlPerChr = 5)
SP2$setSexes(sexes = "yes_rand")
```

***Create new populations***
```{r}
#create population
pop <- newPop(ms_lines, simParam = SP)
#optional step:if the random assignment is either male or female this can change it
#pop@sex[3] <- "F" #Change the sex of the last individual to female

pop2 <- newPop(diverseFounders, simParam = SP2)
#pop2@sex[4:35] <- "M"
```

***Perform random crosses of ms_lines with founder lines***
```{r}
#set.seed(7)
#F1 <- hybridCross(females = pop, males = pop2, crossPlan = "testcross")
#F1 <- hybridCross(females = pop, males = pop2, crossPlan = crossplan, simParam = SP2)
F1 <- randCross2(pop, pop2, 105, nProgeny = 2)
```

***increase recombination rate***
```{r}
# Adjust recombination ratio
#SP$setRecombRatio(2)  
```

***Self F1 generation and generate 6 more generations by SSD***
***This should represent first GORP CYCLE development***
```{r}
# Generate F2 population from each F1 population
for (gen in 1:7) {
  # Perform selfing for each generation of SSD
  ssdPop <- self(F1, nProgeny = 2)  # one progeny per individual (single seed descent)
  
  # Optionally, track the progress of SSD (e.g., genetic diversity loss)
  # print(paste("Generation", gen, "completed"))
}
```

***Pull segregating sites and visualize ***
```{r}
genodata <- pullSegSiteGeno(F1)
genodata2 <- pullSegSiteGeno(ssdPop)
```

***Visualize Data First Cycle***
```{r}
pca <- prcomp(genodata)

pc1 <- pca$x[, 1]  # Principal Component 1
pc2 <- pca$x[, 2]  # Principal Component 2

# Step 6: Plot PCA of the SSD population (Base R)
plot(
  pc1, pc2,
  xlab = "Principal Component 1",
  ylab = "Principal Component 2",
  main = "PCA of Genotypes for Initial Cross",
  col = "blue",      # Point color
  pch = 16,          # Point shape
  cex = 0.7          # Point size
)
grid()

pca <- prcomp(genodata2)

pc1 <- pca$x[, 1]  # Principal Component 1
pc2 <- pca$x[, 2]  # Principal Component 2

# Step 6: Plot PCA of the SSD population (Base R)
plot(
  pc1, pc2,
  xlab = "Principal Component 1",
  ylab = "Principal Component 2",
  main = "PCA of First GORP Cycle",
  col = "blue",      # Point color
  pch = 16,          # Point shape
  cex = 0.7          # Point size
)
grid()
```


***Perform random crosses among first population***
```{r}
F1_2 <- randCross(ssdPop, nCrosses = 105, nProgeny = 2)
```

***Self F1 of the second GORP cycle initial cross and generate 6 more generations by SSD***
***This should represent second GORP CYCLE development***
```{r}
# Generate F2 population from each F1 population
for (gen in 1:7) {
  # Perform selfing for each generation of SSD
  ssdPop2 <- self(F1_2, nProgeny = 1)  # one progeny per individual (single seed descent)
  
  # Optionally, track the progress of SSD (e.g., genetic diversity loss)
  # print(paste("Generation", gen, "completed"))
}
```


***Pull segregating sites and visualize AGAIN***
```{r}
genodata3 <- pullSegSiteGeno(F1_2)
genodata4 <- pullSegSiteGeno(ssdPop2)
```

***Visualize Data second Cycle***
```{r}
pca <- prcomp(genodata3)

pc1 <- pca$x[, 1]  # Principal Component 1
pc2 <- pca$x[, 2]  # Principal Component 2

# Step 6: Plot PCA of the SSD population (Base R)
plot(
  pc1, pc2,
  xlab = "Principal Component 1",
  ylab = "Principal Component 2",
  main = "PCA of Genotypes for second Cross",
  col = "blue",      # Point color
  pch = 16,          # Point shape
  cex = 0.7          # Point size
)
grid()

pca <- prcomp(genodata4)

pc1 <- pca$x[, 1]  # Principal Component 1
pc2 <- pca$x[, 2]  # Principal Component 2

# Step 6: Plot PCA of the SSD population (Base R)
plot(
  pc1, pc2,
  xlab = "Principal Component 1",
  ylab = "Principal Component 2",
  main = "PCA of Second GORP Cycle",
  col = "blue",      # Point color
  pch = 16,          # Point shape
  cex = 0.7          # Point size
)
grid()
```

***Perform random crosses among second population***
```{r}
F1_3 <- randCross(ssdPop2, nCrosses = 210, nProgeny = 2)
```

***Self F1 of the second GORP cycle initial cross and generate 6 more generations by SSD***
***This should represent third GORP CYCLE development***
```{r}
# Generate F2 population from each F1 population
for (gen in 1:7) {
  # Perform selfing for each generation of SSD
  ssdPop3 <- self(F1_3, nProgeny = 1)  # one progeny per individual (single seed descent)
}
```


***Pull segregating sites and visualize AGAIN x3***
```{r}
genodata5 <- pullSegSiteGeno(F1_3)
genodata6 <- pullSegSiteGeno(ssdPop3)
```

***Visualize Data second Cycle***
```{r}
pca <- prcomp(genodata5)

pc1 <- pca$x[, 1]  # Principal Component 1
pc2 <- pca$x[, 2]  # Principal Component 2

# Step 6: Plot PCA of the SSD population (Base R)
plot(
  pc1, pc2,
  xlab = "Principal Component 1",
  ylab = "Principal Component 2",
  main = "PCA of Genotypes for third Cross",
  col = "blue",      # Point color
  pch = 16,          # Point shape
  cex = 0.7          # Point size
)
grid()

pca <- prcomp(genodata6)

pc1 <- pca$x[, 1]  # Principal Component 1
pc2 <- pca$x[, 2]  # Principal Component 2

# Step 6: Plot PCA of the SSD population (Base R)
plot(
  pc1, pc2,
  xlab = "Principal Component 1",
  ylab = "Principal Component 2",
  main = "PCA of Third GORP Cycle",
  col = "blue",      # Point color
  pch = 16,          # Point shape
  cex = 0.7          # Point size
)
grid()
```

***Perform random crosses among third population***
```{r}
F1_4 <- randCross(ssdPop3, nCrosses = 420, nProgeny = 2)
```

***Self F1 of the second GORP cycle initial cross and generate 6 more generations by SSD***
***This should represent third GORP CYCLE development***
```{r}
# Generate F2 population from each F1 population
for (gen in 1:7) {
  # Perform selfing for each generation of SSD
  ssdPop4 <- self(F1_4, nProgeny = 1)  # one progeny per individual (single seed descent)
}
```


***Pull segregating sites and visualize AGAIN x3***
```{r}
genodata7 <- pullSegSiteGeno(F1_4)
genodata8 <- pullSegSiteGeno(ssdPop4)
```

***Visualize Data second Cycle***
```{r}
pca <- prcomp(genodata7)

pc1 <- pca$x[, 1]  # Principal Component 1
pc2 <- pca$x[, 2]  # Principal Component 2

# Step 6: Plot PCA of the SSD population (Base R)
plot(
  pc1, pc2,
  xlab = "Principal Component 1",
  ylab = "Principal Component 2",
  main = "PCA of Genotypes for fourth Cross",
  col = "blue",      # Point color
  pch = 16,          # Point shape
  cex = 0.7          # Point size
)
grid()

pca <- prcomp(genodata8)

pc1 <- pca$x[, 1]  # Principal Component 1
pc2 <- pca$x[, 2]  # Principal Component 2

# Step 6: Plot PCA of the SSD population (Base R)
plot(
  pc1, pc2,
  xlab = "Principal Component 1",
  ylab = "Principal Component 2",
  main = "PCA of Fourth GORP Cycle",
  col = "blue",      # Point color
  pch = 16,          # Point shape
  cex = 0.7          # Point size
)
grid()
```

